---
title: "Chapter 2 labs"
author: "Russ Conte"
date: "8/6/2021"
output: html_document
---

# 2.3.1 Basic Commands


```{r Chapter 2 labs, include=FALSE}
x <-  c(1,3,2, 5) # a basic vector
x # returns the vector

x <-  c(1, 6, 2)
x

y <- c(1, 4,3)
y # returns y
```

## Adding two sets of numbers together

From the text:

We can tell R to add two sets of numbers together. It will then add the first number from x to the first number from y, and so on. However, x and y should be the same length. We can check their length using the length() function.

```{r Adding two sets of numbers together}
length(x)
length(y)
# note length(x) = length(y)
x + y 

```

## ls() allows us to look at a list of ALL of the objects we have saved so far


```{r ls(), echo=FALSE}
ls()
rm(x,y) # removes specified items
ls() # checks which objects we have - note we don't have any 

```

## Making a matrix
```{r Making a matrix}
x = matrix(data = c(1,2,3,4), nrow = 2, ncol = 2)
x

# same idea, done a different way:
x = matrix(c(1,2,3,4),2,2)
x

# note that it is possible to assign by rows or columns
matrix(c(1,2,3,4),2,2,byrow = TRUE)

# note that it's possible to apply functions, such as sqrt and ^2 to matrices

sqrt(x)
x^2
x^2.7182818
```

```{r, rnorm function}
x=rnorm(50) # will generate 50 random normal values
y = x + rnorm(50, mean = 50, sd = 0.1)
cor(x,y)

```

## to create reproducible code, we can set the seed
```{r, Set Seed}
set.seed(1303)
rnorm(50)


```

## using mean and var with a vector of numbers

```{r, mean and var with a vector of numbers}
set.seed(3)
y = rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)

```

## 2.3.2 Graphics