---
title: "Chapter 2 labs"
author: "Russ Conte"
date: "8/6/2021"
output: html_document
---

# 2.3.1 Basic Commands


```{r Chapter 2 labs, include=FALSE}
x <-  c(1,3,2, 5) # a basic vector
x # returns the vector

x <-  c(1, 6, 2)
x

y <- c(1, 4,3)
y # returns y
```

## Adding two sets of numbers together

From the text:

We can tell R to add two sets of numbers together. It will then add the first number from x to the first number from y, and so on. However, x and y should be the same length. We can check their length using the length() function.

```{r Adding two sets of numbers together}
length(x)
length(y)
# note length(x) = length(y)
x + y 

```

## ls() allows us to look at a list of ALL of the objects we have saved so far


```{r ls(), echo=FALSE}
ls()
rm(x,y) # removes specified items
ls() # checks which objects we have - note we don't have any 

```

## Making a matrix
```{r Making a matrix}
x = matrix(data = c(1,2,3,4), nrow = 2, ncol = 2)
x

# same idea, done a different way:
x = matrix(c(1,2,3,4),2,2)
x

# note that it is possible to assign by rows or columns
matrix(c(1,2,3,4),2,2,byrow = TRUE)

# note that it's possible to apply functions, such as sqrt and ^2 to matrices

sqrt(x)
x^2
x^2.7182818
```

```{r, rnorm function}
x=rnorm(50) # will generate 50 random normal values
y = x + rnorm(50, mean = 50, sd = 0.1)
cor(x,y)

```

## to create reproducible code, we can set the seed
```{r, Set Seed}
set.seed(1303)
rnorm(50)


```

## using mean and var with a vector of numbers

```{r, mean and var with a vector of numbers}
set.seed(3)
y = rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)

```

## 2.3.2 Graphics

From the text: The plot() function is the primary way to plot data in R

```{r, plotting examples}
x = rnorm(100)
y = rnorm(100)
plot(x,y)
plot(x, y, xlab = "This is the x-axis", ylab = "This is the y-axis", main = "Plot of x versus y")


```

## creating a sequence of numbers using seq:

```{r, creating a sequence of numbers}

x = seq(1,10)
x
x = 1:10 # alternate way to accomplish the same goal
x

x = seq(-pi, pi, length = 50) # creates 50 values between -pi and pi
x
```

## create contour plots.

From the text:
The contour() function produces a contour plot in order to represent three-dimensional data; it is like a topographical map.

```{r, contour plots}

y = x
f = outer(x, y, function (x,y) cos(y)/(1+x^2))
contour(x,y,f)
fa = (f - t(f)) / 2
contour(x, y, fa, nlevels = 15)

```

## image function

from the text: The image() function works the same way as contour(), except that it produces a color-coded plot whose colors depend on the z value.

```{r, image function}

image(x, y, fa)
persp(x, y, fa)
persp(x, y, fa, theta = 30)
persp(x, y, fa, theta = 30, phi = 20)
persp(x, y, fa, theta = 30, phi = 40)

```

## 2.3.3 Indexing Data

From the text:
We often wish to examine part of a set of data. Suppose that our data is stored in the matrix A.

```{r, indexing data}
A = matrix(1:16, 4,4)
A


```

Then typing A[2,3] will return the item in the 2nd row and 3rd column   

```{r}
A[2,3]
```

From the text: We can also select multiple rows and columns at a time, by providing vectors as the indices.

```{r, selecting multiple rows and columns by indexing}

A[c(1,3), c(2,4)]

A[1:3, 2:4] # rows 1, 2, and 3, and columns 2, 3, and 4

A[1:2,] # first two rows, all the columns

A[,1:2] # all the rows, the first two columns

A[1,] # the first row, all the columns



```

From the text: The use of a negative sign - in the index tells R to keep all rows or columns except those indicated in the index.

```{r, using a negative index}

A[-c(1,3),] # not rows 1 and 3, include all columns

A[-c(1,3), -c(1,3,4)]

dim(A) # gives the dimensions of a matrix or data frame

```

## 2.3.5 Additional Graphical and Numerical Summaries

From the text: To refer to a variable, we must type the data set and the variable name joined with a $ symbol. 

```{r}
Auto = ISLR::Auto
View(Auto)
plot(Auto$cylinders, Auto$mpg)

```

From the text: There are various ways to deal with the missing data. In this case, only five of the rows contain missing observations, and so we choose to use the na.omit() function to simply remove these rows.

```{r, dealing with missing data}

Auto = na.omit(Auto)
dim(Auto)

```

## 2.3.5 Additional Graphical and Numerical Summaries

```{r, additional summaries}

plot(Auto$cylinders, Auto$mpg)
plot(cylinders, mpg)
cylinders = as.factor(cylinders)

plot(cylinders, mpg)
plot(cylinders, mpg, col = "red")
plot(cylinders, mpg, col = "red", varwidth = T)
plot(cylinders, mpg, col = "red", varwidth = T, horizontal = T)
plot(cylinders, mpg, col = "red", varwidth = T, xlab = "cylinders", ylab = "MPG")



```

From the text: The hist() function can be used to plot a histogram. Note that col=2 has the same effect as col="red".

```{r, plotting histograms}

hist(mpg)
hist(mpg, col = 2)
hist(mpg, col = 2, breaks = 15)

```

From the text: The pairs() function creates a scatterplot matrix i.e. a scatterplot for every pair of variables for any given data set. We can also produce scatterplots for just a subset of the variables.

```{r, pairs scatterplots}

pairs(Auto)
pairs(~mpg + displacement + horsepower + weight + acceleration, Auto)

```

From the text:
The summary() function produces a numerical summary of each variable in a particular data set.

```{r, summary}
summary(Auto)
summary(mpg)
```

