---
title: "Chapter 2 labs"
author: "Russ Conte"
date: "8/6/2021"
output: html_document
---

# 2.3.1 Basic Commands


```{r Chapter 2 labs, include=FALSE}
x <-  c(1,3,2, 5) # a basic vector
x # returns the vector

x <-  c(1, 6, 2)
x

y <- c(1, 4,3)
y # returns y
```

## Adding two sets of numbers together

From the text:

We can tell R to add two sets of numbers together. It will then add the first number from x to the first number from y, and so on. However, x and y should be the same length. We can check their length using the length() function.

```{r Adding two sets of numbers together}
length(x)
length(y)
# note length(x) = length(y)
x + y 

```

## ls() allows us to look at a list of ALL of the objects we have saved so far


```{r ls(), echo=FALSE}
ls()
rm(x,y) # removes specified items
ls() # checks which objects we have - note we don't have any 

```

## Making a matrix
```{r Making a matrix}
x = matrix(data = c(1,2,3,4), nrow = 2, ncol = 2)
x

# same idea, done a different way:
x = matrix(c(1,2,3,4),2,2)
x

# note that it is possible to assign by rows or columns
matrix(c(1,2,3,4),2,2,byrow = TRUE)

# note that it's possible to apply functions, such as sqrt and ^2 to matrices

sqrt(x)
x^2
x^2.7182818
```

```{r, rnorm function}
x=rnorm(50) # will generate 50 random normal values
y = x + rnorm(50, mean = 50, sd = 0.1)
cor(x,y)

```

## to create reproducible code, we can set the seed
```{r, Set Seed}
set.seed(1303)
rnorm(50)


```

## using mean and var with a vector of numbers

```{r, mean and var with a vector of numbers}
set.seed(3)
y = rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)

```

## 2.3.2 Graphics

From the text: The plot() function is the primary way to plot data in R

```{r, plotting examples}
x = rnorm(100)
y = rnorm(100)
plot(x,y)
plot(x, y, xlab = "This is the x-axis", ylab = "This is the y-axis", main = "Plot of x versus y")


```

## creating a sequence of numbers using seq:

```{r, creating a sequence of numbers}

x = seq(1,10)
x
x = 1:10 # alternate way to accomplish the same goal
x

x = seq(-pi, pi, length = 50) # creates 50 values between -pi and pi
x
```

## create contour plots.

From the text:
The contour() function produces a contour plot in order to represent three-dimensional data; it is like a topographical map.

```{r, contour plots}

y = x
f = outer(x, y, function (x,y) cos(y)/(1+x^2))
contour(x,y,f)
fa = (f - t(f)) / 2
contour(x, y, fa, nlevels = 15)

```

## image function

from the text: The image() function works the same way as contour(), except that it produces a color-coded plot whose colors depend on the z value.

```{r, image function}

image(x, y, fa)
persp(x, y, fa)
persp(x, y, fa, theta = 30)
persp(x, y, fa, theta = 30, phi = 20)
persp(x, y, fa, theta = 30, phi = 40)

```

